---
title: "What Determines House Prices in King County, USA?"
author: "Mariam Badmus"
date: "26-10-2022"
output: 
 html_document:
  toc_depth: 3

---

\newpage 
\tableofcontents 
\newpage



### 1.   Read the data

```{r}
kc_house_data <- read.csv("kc_house_data.csv")
```


```{r load kc_house_data, include=TRUE, message= FALSE, warning=FALSE, echo = TRUE, eval= FALSE}

load("C:/Users/kike_/OneDrive/Documents/house_prediction/predicting_house_price_in_kingcounty_usa.RData")
```




### 2.     Install Packages

```{r}
library(psych)
library(graphics)
library(corrplot)
library(ggplot2)
library(tidyr)
library(reshape2)
library(dplyr)
```




### 3.     Inspecting data 
```{r}
glimpse(kc_house_data)
```




### 4.     Checking for duplicates: we have zero duplicate  in our data 

```{r}
sum(duplicated(kc_house_data)) 

```




### 5.     Check the relationships between price and other variables

Using library(psych) to create scatter plot to pinpoint possible relationships between price and other variables. According to the scatterplots below, only sqft livng and sqft above have a relationship  with price, which make them suitable for a  regression. 


```{r}
 kc_melt <- melt(kc_house_data, id.vars = "price", measure.vars = c("sqft_living", "grade" , "bathrooms" , "sqft_above", "sqft_living15", "sqft_living",
  "bedrooms", "floors", "sqft_basement", "lat", "view", "condition", "yr_built", "yr_renovated", "long", "sqft_lot15" , "zipcode" ,  "sqft_lot"))


ggplot(kc_melt,(aes(x = value, y = price)))+
  geom_point()+
  geom_smooth ()+
  facet_wrap(vars(variable), scales= "free")+
  ggtitle('Graph 1: Scatterplots of Price against Other variables')


ggplot(kc_house_data[,3:21], aes(x = sqft_living, y = price))+ #visualising sqft_living independently
  geom_smooth(aes())+
   scale_y_continuous(limits = c(0, 8000000), breaks = seq(0, 8000000, by = 500000), labels = scales::comma)+
   ggtitle('Graph 2: Scatterplots of Price Sqft_Living')


ggplot(kc_house_data[,3:21], aes(x = sqft_above, y = price))+ 
  geom_smooth(aes())+
   scale_y_continuous(limits = c(0, 8000000), breaks = seq(0, 8000000, by = 500000), labels = scales::comma)+
    ggtitle('Graph 3: Scatterplots of Price Sqft_Above')
```





### 6.     Finding the corrrelation between  variables from column 3 and column 21 and multicolinearity amongst the other variables with texts

This data set has more than three independent variables set so it will be difficult to determine if all variables are correlated with 2d or 3d plots.

```{r}

correlation <- cor(kc_house_data[,3:21])
levels <- colnames(kc_house_data[,3:21])

 correlation %>% 
  data.frame() %>% 
  mutate(row = factor(rownames(.), levels = levels),  rowid  = as.numeric(row)) %>% #as factor level by col_names in levels
  pivot_longer(-c(row, rowid), names_to = "col") %>% #row and rowid as column...col as 2nd column
  mutate(col = factor(col, levels = levels),  colid  = as.numeric(col)) %>% 
  ggplot2::ggplot(aes(col,row))+
  geom_text(aes(label = scales::number(value, accuracy = .01), color = abs(value)), #access number from scale function #round value to two decimal digits
            size = 2.5) +
  scale_x_discrete(labels = ~ attr(.x, "pos"), drop = FALSE) + #setting the value of x independent of no variable
  scale_y_discrete(labels = ~ paste0(.x, " (", attr(.x, "pos"), ")"), drop = FALSE)+ #concatenate y axis with brackets with x independent of no variable
     ggtitle('Graph 4: Correlation Matrix for All Variables')
 
corrplot(cor(kc_house_data[,3:21]), method = "circle")
```


### 7.     Using lm function to build our linear regression model using variables where correlation >= 0.7 and visualising 


```{r}
price_sqftliving <- lm(price ~ sqft_living , data = kc_house_data)

summary(price_sqftliving)

plot(price_sqftliving)

abline(price_sqftliving) 



```

Living space has the strongest correlation with prices at r = 0.70



### 8.    Using the moderate correlated variables and use it to build a prediction model 


```{r}
living_above <- lm(price ~ sqft_living, data = kc_house_data)
  
summary(living_above)

plot(living_above)

abline(living_above)
```


### 9.    Add all the Variables

```{r}
price_alldata <- lm(price ~ sqft_living + grade  + bathrooms + sqft_above +sqft_living15 + sqft_living + 
  bedrooms + floors + sqft_basement + lat +  view + condition + yr_built + yr_renovated + long + sqft_lot15 
  + zipcode + sqft_lot, data = kc_house_data)

summary(price_alldata)

plot(price_alldata)

abline(price_alldata)
```





### 10.    Test price y within the range of the three multiple regression built using R squared Adjusted; compare it to the observed data
Simple linear regression is y = b + mx #b = intercept, x = slope and the other varibales with multiple regression using y = y = b1x1 + b2x2 + … + bnxn + c.

```{r}
#sqft_living- let give x = 1960 in observation 4
-43580.743 + 280.624 * 1960

#for price against  sqft_living sqft_above 
 -40859.96 +  (295.50 * 1960) + (-18.83 * 1050)

#for price against
-7.522e+06 + (1.505e+02  * 1960) + (9.414e+04 * 7) + (4.061e+04 * 3.00) + (3.645e+01 * 1050) + (1.835e+0 * 1360) + (-3.862e+04 * 4) + (7.099e+03 * 1.0) + (5.998e+05  * 47.5208) + (7.980e+04 * 0) + (2.720e+04 * 5) + (-2.590e+03  * 1965) + (2.663e+01  * 0) + (-2.229e+05 * -122.393) + ( -3.543e-01 * 5000) + (-6.001e+02 * 98136) + (1.010e-01  * 5000)
```






### 12.   Predict y with a random x variable

Prediciting house price with sqft_lving as an independent variable provides a price close to the actual price value in the data, adding sqft_above increased the price but not by a lot but still within the price in obseravtion 4 of the main data. It si best to leave out sqft_above given the high correlation fo both data at 0.88.I will be plotting the regression graph using the intercept and slope given for value of price and sqft_living. 


```{r}

-43580.743 + 280.624 * 3000 #using a randome x variable of 3000 sqft_living then compare the results to graph 6 below

ggplot2::ggplot(kc_house_data[,3:21], aes(x = sqft_living, y = price))+ #visualising sqft_living independently
  geom_smooth(aes())+
   scale_y_continuous(limits = c(0, 8000000), breaks = seq(0, 8000000, by = 500000), labels = scales::comma)+
  scale_x_continuous(limits = c(0, 14000), breaks = seq(0, 14000, by = 2000), labels = scales::comma)+
    ggtitle('Graph 5: Compare Predicted Value for Y with Actual Data Points')



xs <- kc_house_data[,6]
ys = kc_house_data[,3]

plot(xs,ys)
abline(a = -43580.743, b = 280.624)
```





### 13.    Conclusion

The y intercept for price(dependent variable) is negative when x(independent variable) is which does not apply to practical because a price cannot be negative without a living space. Using multiple regression have proven to be poorly suited to predict house prices in King county, USA using other variables outside sqft_living. 
















